outlier_exploration

### 2025-05-07 Kyubin's Edit/KYUBIN_EDIT.qmd
Lines 96-194 were edited: Added in the Code taken from "multicolinearity_and_outlier.qmd" that justifies obs 264 is an outlier Created the new dataset that removed obs 264 called sleep_test Created the new full model that removed obs 264 called full_model_test Changed "sleep" to "sleep_test" for for the VIF testing

@@

vif(full_model)

#GVIF high for occupation, stress, bmi category
#GVIF^ high for quality of sleep

#stress
sleep$stress_level_num <- as.numeric(as.factor(sleep$stress_level))

aux_model_stress_level <- lm(stress_level_num ~ gender + age + quality_of_sleep + 
                    physical_activity_level + occupation + 
                    bmi_category + heart_rate + daily_steps + 
                    sleep_disorder + MAP, data = sleep)
summary(aux_model_stress_level)
# R^2 .9542 suggest we remove it

#occupation
sleep$occupation_num <- as.numeric(as.factor(sleep$occupation))

aux_model_occupation <- lm(occupation_num ~ gender + age + quality_of_sleep + 
                           physical_activity_level + bmi_category + heart_rate + 
                           daily_steps + sleep_disorder + MAP + stress_level, 
                           data = sleep)

summary(aux_model_occupation)
# R^2 around .7232 keep occupation

# BMI
sleep$bmi_category_num <- as.numeric(as.factor(sleep$bmi_category))
aux_model_bmi <- lm(bmi_category_num ~ gender + age + quality_of_sleep + 
                     physical_activity_level + occupation + heart_rate + 
                     daily_steps + sleep_disorder + MAP + stress_level,
                     data = sleep)

summary(aux_model_bmi)
# R^2 around 0.9375

#quality
sleep$quality_of_sleep_num <- as.numeric(as.factor(sleep$quality_of_sleep))
aux_model_quality <- lm(quality_of_sleep_num ~ gender + age + physical_activity_level +
                         occupation + bmi_category + heart_rate + daily_steps +
                         sleep_disorder + MAP + stress_level,
                         data = sleep)

summary(aux_model_quality)
# R^2 around 0.9653

>>>>

### Code taken from "multicolinearity_and_outlier.qmd" that justifies obs 264 is an outlier

```{r}
leverage <- hatvalues(full_model)
studentized <- rstudent(full_model)
dffits_vals <- dffits(full_model)
cooks_d <- cooks.distance(full_model)

diagnostics <- data.frame(Leverage = leverage,
                          Studentized = studentized,
                          DFFITS = dffits_vals,
                          Cooks_Distance = cooks_d)

n <- nrow(sleep)
p <- length(coef(full_model)) - 1
leverage_threshold <- 2 * (p + 1) / n
studentized_threshold <- 2
cooks_threshold <- 1

# Find points exceeding thresholds
high_leverage <- which(leverage > leverage_threshold)
high_studentized <- which(abs(studentized) > studentized_threshold)
high_cooks <- which(cooks_d > cooks_threshold)

# Unique indices that exceed any threshold
suspicious_indices <- sort(unique(c(high_leverage, high_studentized, high_cooks)))

# Return threshold values, diagnostic preview, and rows flagged as potential outliers
list(
  leverage_threshold = leverage_threshold,
  studentized_threshold = studentized_threshold,
  cooks_threshold = cooks_threshold,
  diagnostics_preview = head(diagnostics, 10),
  flagged_rows = diagnostics[suspicious_indices, ])

# observation 264 has a leverage of 1

# New data set used for model
sleep_test <- sleep[-264,]

full_model_test<-lm(sleep_duration ~., sleep_test)
summary(full_model_test)
```

@@

vif(full_model_test)

#GVIF high for occupation, stress level, bmi category
#GVIF^ high for quality of sleep

#stress
sleep_test$stress_level_num <- as.numeric(as.factor(sleep_test$stress_level))

aux_model_stress_level <- lm(stress_level_num ~ gender + age + quality_of_sleep + 
                    physical_activity_level + occupation + 
                    bmi_category + heart_rate + daily_steps + 
                    sleep_disorder + PP, data = sleep_test)
summary(aux_model_stress_level)
# R^2 .9563 suggest we remove it

#occupation
sleep_test$occupation_num <- as.numeric(as.factor(sleep_test$occupation))

aux_model_occupation <- lm(occupation_num ~ gender + age + quality_of_sleep + 
                           physical_activity_level + bmi_category + heart_rate + 
                           daily_steps + sleep_disorder + PP + stress_level, 
                           data = sleep_test)

summary(aux_model_occupation)
# R^2 around .7453, keep occupation

# BMI
sleep_test$bmi_category_num <- as.numeric(as.factor(sleep_test$bmi_category))

aux_model_bmi <- lm(bmi_category_num ~ gender + age + quality_of_sleep + 
                     physical_activity_level + occupation + heart_rate + 
                     daily_steps + sleep_disorder + PP + stress_level,
                     data = sleep_test)

summary(aux_model_bmi)
# R^2 around 0.9326

#quality
sleep_test$quality_of_sleep_num <- as.numeric(as.factor(sleep_test$quality_of_sleep))

aux_model_quality <- lm(quality_of_sleep_num ~ gender + age + physical_activity_level +
                         occupation + bmi_category + heart_rate + daily_steps +
                         sleep_disorder + PP + stress_level,
                         data = sleep_test)

summary(aux_model_quality)
# R^2 around 0.9673
